---
title: "Assignment2"
author: "Louis"
date: "2022-10-05"
output: html_document
---
# Assignment2
```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```
## 1.1(Q1)
```{r}
# hSF <- select(filter(Hawks, Species == "RT" & Weight >= 1000), Wing, Weight, Tail)
hSF <- Hawks %>% filter(Species == "RT" & Weight >= 1000) %>% select(Wing, Weight, Tail)
#hSF
```
## 1.1(Q2)
3 variables
398 examples, observations or cases
## 1.2(Q1)
```{r}
head(hSF %>% arrange(Wing), 5)
```
## 1.3(Q1)
```{r}
hawkSpeciesNameCodes <- data.frame(species_code = c("CH", "RT", "SS"), species_name_full = c(" Cooper's", "Red-tailed", " Sharp-shinned"))
```
## 1.3(Q2)
```{r}
hawksFullName <- Hawks %>% left_join(rename(hawkSpeciesNameCodes, Species = species_code)) %>% select(-Species) %>% rename(Species = species_name_full)
```
## 1.3(Q3)
```{r}
head(select(hawksFullName, Species, Wing, Weight), 7)
```
In this situation there is no difference between different types of join function because the data frame called "hawkSpeciesNameCodes" contains all three species in "Hawks" and it doesn't include extra species.  
When there is more species in any data frame or "hawkSpeciesNameCodes" contains fewer species, using different join functions would make a difference.

## 1.4(Q1)
```{r}
hawksWithBMI <- Hawks %>% mutate(bird_BMI = 1000 * Weight / Wing ^ 2) %>% select(Species, bird_BMI) %>% arrange(desc(bird_BMI))
head(hawksWithBMI, 8)
```
## 1.5(Q1)
```{r}
summary_table <- hawksFullName %>% group_by(Species) %>% summarize(num_rolls = n(), mn_wing = mean(Wing), nd_wing = median(Wing, na.rm = TRUE), t_mn_wing = mean(Wing, trim = 0.1), b_wt_ratio = max(Wing/Tail, na.rm = TRUE))
#when to make na.rm TRUE?
summary_table
```
## 1.5(Q2)
```{r}
missing_values <- hawksFullName %>% group_by(Species) %>% select(Species, Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Crop) %>% summarize(across(everything(), ~ sum(is.na(.))))
missing_values
```

## 2.1(Q2)
```{r}
impute_by_median<-function(x){
  mu<-median(x,na.rm=1)
  impute_f<-function(z){
    if(is.na(z)){
      return(mu)
    }else{
      return(z)
    } }
  return(map_dbl(x,impute_f))
}
v<-c(1,2,NA,4)
impute_by_median(v)
```
## 2.1(Q3)
```{r}
x <- seq(0, 10, 0.1)
#df_xy <- data.frame(x, y = 5 * x + 1)
fun1 <- function(x){
  return (5 * x + 1)
}
fun2 <- function(x, y){
  return (x + y)
}
y = map_dbl(x, fun1)
df_xy <-data.frame(x, y)
df_xy %>% head(5)
df_xyz <- data.frame(x, y, z = map2_dbl(x, y, fun2))
df_xyz %>% head(5)
```
## 2.1(Q4)
```{r}
sometimes_missing <- function(index, value){
  if(index %% 5 == 0){
    return (NA)
  }
  else{
    return (value)
  }
}
sometimes_missing(14,25)
sometimes_missing(15,25)
df_xy_missing <- data.frame(x, y = map2_dbl(row_number(x), map_dbl(x, fun1), sometimes_missing))
df_xy_missing %>% head(10)
```
## 2.1(Q5)
```{r}
df_xy_imputed <- df_xy_missing %>% mutate(z = impute_by_median(y)) %>% select(-y) %>% rename(y = z)
df_xy_imputed %>% head(6)
```

```{r}

```