---
title: "Assignment 3"
author: "Louis Wang"
date: "2022-10-12"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# Load packages
library(tidyverse)
library(Stat2Data)
data("Hawks")
```
# 1. Exploratory data analysis
```{r}
head(Hawks)
```

## 1.1 (Q1)   
```{r}
# the Tail column of Hawks data frame
HawksTail <- Hawks[, "Tail"]
# compute the sample mean and sample median
meanTail <- mean(HawksTail)
medianTail <- median(HawksTail)
```

## 1.2 (Q1)
```{r}
#  compute the sample mean, sample median and trimmed sample mean (with q = 0.5) of the Hawk’s wing length and Hawk’s weight
sampleSummary <- Hawks %>% summarise(Wing_mean = mean(Wing, na.rm = TRUE),
                                     Wing_t_mean = mean(Wing, trim = 0.5, na.rm = TRUE),
                                     Wing_med = median(Wing, na.rm = TRUE),
                                     Weight_mean = mean(Weight, na.rm = TRUE),
                                     Weight_t_mean = mean(Weight, trim = 0.5, na.rm = TRUE),
                                     Weight_med = median(Weight, na.rm = TRUE))
sampleSummary
```

## 1.2 (Q2)
```{r}
# Combine them with the group_by() function to obtain a breakdown by species
sampleSummaryBySpecies <- Hawks %>% group_by(Species) %>% summarise(Wing_mean = mean(Wing, na.rm = TRUE),
                                                                    Wing_t_mean = mean(Wing, trim = 0.5, na.rm = TRUE),
                                                                    Wing_med = median(Wing, na.rm = TRUE),
                                                                    Weight_mean = mean(Weight, na.rm = TRUE),
                                                                    Weight_t_mean = mean(Weight, trim = 0.5, na.rm = TRUE),
                                                                    Weight_med = median(Weight, na.rm = TRUE))
sampleSummaryBySpecies
```

## 1.3 (Q1)
```{r}
#  letting a = 2 and b = 3, compute the mean of HawksTail*a+b
fun1 <- function(x){
  return (2 * x + 3)
}
anotherTail <- map_dbl(HawksTail, fun1)
meanAnotherTail <- mean(anotherTail)
# and then compare it with the one obtained from the mean of HawksTail
meanAnotherTail == meanTail * 2 + 3
```

## 1.3 (Q2)
```{r}
# variance and standard deviation 
varTail <- var(HawksTail, na.rm = TRUE)
sdTail <- sd(HawksTail, na.rm = TRUE)
varAnotherTail <- var(anotherTail, na.rm = TRUE)
sdAnotherTail <- sd(anotherTail, na.rm = TRUE)
# compare
varAnotherTail == 4 * varTail & sdAnotherTail == 2 * sdTail
```

## 1.4
```{r}
# a vector called “hal” consisting of the talon lengths of all the hawks with any missing values removed
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans
# generate a new vector called “corrupted_hall” with 10 outliers each of value 100
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
#  compute the mean of the original sample and the corrupted sample
mean(hal)
mean(corrupted_hal)
# generates a vector called “means_vect” which gives the sample means of corrupted samples with different numbers of outliers
# means_vect is a vector of length 1001 with the i-th entry equal to the mean of a sample with i − 1 outliers
num_outliers_vect <- seq(0,1000)
means_vect <- c()
for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
  means_vect <- c(means_vect, mean(corrupted_hal))
}
```

## 1.4 (Q1)
```{r}
# Sample median
medians_vect <- c()
for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
  medians_vect <- c(medians_vect, median(corrupted_hal))
}
```

## 1.4 (Q2)
```{r}
# Sample trimmed mean
t_means_vect <- c()
for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
  t_means_vect <- c(t_means_vect, mean(corrupted_hal, trim = 0.1))
}
```

## 1.4 (Q3)
```{r}
# Visualisation
df_means_medians <- data.frame(num_outliers=num_outliers_vect, mean=means_vect, t_mean=t_means_vect, median=medians_vect)
df_means_medians %>% pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value") %>%
                     ggplot(aes(x=num_outliers,color=Estimator, linetype=Estimator,y=Value)) + geom_line()+xlab("Number of outliers")
```

## 1.5 (Q1)
```{r}
#  create a box plot which summarises the distribution of hawk weights broken down by species
ggplot(data = Hawks, aes(x = Species, y = Weight)) + geom_boxplot()
```

## 1.5 (Q2)
```{r}
# quantile and boxplots
weightQuentile <- Hawks %>% group_by(Species) %>% 
  summarise(quantile025 = quantile(Weight, prob = 0.25, na.rm = TRUE), 
            quantile050 = quantile(Weight, prob = 0.5, na.rm = TRUE),
            quantile075 = quantile(Weight, prob = 0.75, na.rm = TRUE))
weightQuentile
```

## 1.5 (Q3)
```{r}
# outliers
num_outliers <- function(x){
  x<-x[!is.na(x)]
  q25 <- quantile(x, prob = 0.25)
  q75 <- quantile(x, prob = 0.75)
  IQX <- q75 - q25
  result <- length(x[(x > q75 + 1.5 * IQX) | (x < q25 - 1.5 * IQX)])
  return (result)
}
num_outliers( c(0, 40,60,185))
```

## 1.5 (Q4)
```{r}
# Outliers by group
Outliers <- Hawks %>% group_by(Species) %>% 
  summarise(num_outliers_weight = num_outliers(Weight))
Outliers
```

## 1.6 (Q1)
```{r}
# Compute the covariance and correlation between the Weight and Wing of the Hawks data
covariance <- cov(Hawks$Weight, Hawks$Wing, use = 'complete.obs')
covariance
correlation <- cor(Hawks$Weight, Hawks$Wing, use = 'complete.obs')
correlation
```

## 1.6 (Q2)
```{r}
# a = 2.4, b = 7.1, c = −1, d = 3
fun2 <- function(x){
  return (2.4 * x + 7.1)
}
fun3 <- function(x){
  return (-1 * x + 3)
}
newCovariance <- cov(fun2(Hawks$Weight), fun3(Hawks$Wing), use = 'complete.obs')
newCorrelation <- cor(fun2(Hawks$Weight), fun3(Hawks$Wing), use = 'complete.obs')
# What is the covariance between a * Weight + b and c * Wing + d ?
# What is the correlation between a * Weight + b and c * Wing + d ?
newCovariance == -2.4 * covariance & (newCorrelation - correlation * -1) < 1e-10

```

# 2. Random experiments, events and sample spaces, and the set theory

## 2.1 (Q1)
- random experiment is a procedure which has a well-defined set of possible outcomes **and** could (at least in principle) be repeated arbitrarily many times
- event is a set of possible outcomes of an experiment
- sample space is the set of all possible outcomes of interest for a random experiment

## 2.1 (Q2)
- 2 of {1,2,3,4,5,6}
- {{1,1},{1,2},{1,3},{1,4},{1,5},{1,6},  
{2,1},{2,2},{2,3},{2,4},{2,5},{2,6},  
{3,1},{3,2},{3,3},{3,4},{3,5},{3,6},  
{4,1},{4,2},{4,3},{4,4},{4,5},{4,6},  
{5,1},{5,2},{5,3},{5,4},{5,5},{5,6},  
{6,1},{6,2},{6,3},{6,4},{6,5},{6,6}}
- total: 36
- no

## 2.2 (Q1)
1. A ∪ B = {1,2,3,4,6}, A ∪ C = {1,2,3,4,5,6}
2. A ∩ B = {2}, A ∩ C = ∅
3. A \ B = {1,3},  A \ C = {1,2,3}
4. no, yes
5. yes
6. {1,2,3}, {4,5,6}    {1,2}, {3,4}, {5,6}

## 2.2 (Q2)
1. A
2. ∅
3. 

## 2.2 (Q3)
cardinality = 2 ^ K

## 2.2 (Q4)
1. ∅
2. S1∩S2=∅, S1∪S2=A1∪A2, S1∩S2∩S3=∅, S1∪S2∪S3=A1∪A2∪A3, S1∩S2∩S3∩S4=∅, S1∪S2∪S3∪S4=A1∪A2∪A3∪A4

## 2.2 (Q5)
1. 1Ac = 1 - 1A
2. Ω
3. 

# 3. Visualisation

## 3 (Q1)
```{r}
#  Density plot
ggplot(data = Hawks, aes(x = Tail, color = Species)) + geom_density() + xlab("Tail(mm)")
```

## 3 (Q2)
```{r}
#  Violin plot
ggplot(data = Hawks, aes(x = Tail, y = Species, fill = Species)) + geom_violin() + xlab("Tail(mm)")
```

## 3 (Q3)
```{r}
# Scatter plot
scatter <- ggplot(data = Hawks, aes(x = Tail, y = Weight, color = Species)) + xlab("Tail(mm)") + ylab("Weight(mm)")
scatter + geom_point()
```

## 3 (Q4)
```{r}
# Trend lines and facet wraps
scatter + geom_point() + facet_wrap(vars(Species), scales='free') + geom_smooth(method = "lm")
```

## 3(Q5)
```{r}
# Adding annotations
max(filter(Hawks)$Weight, na.rm = TRUE)
scatter + geom_point() + 
  geom_curve(x = 180, xend = 196, y = 1750, yend = 2030, arrow = arrow(length = unit(0.5, 'cm')), curvature = 0.1, color = 'black') + 
  geom_text(x = 175, y = 1900, label = "heaviest hawk", color = 'black')
```




